thesis

### 2024-12-26 thesis.qmd
SFR translation and hubble

## Galaxy Ingridients

@@

### Hubble-de Vacouler Classification

@@

## Star-Formnig Regions

@@

Understanding SFH models is crucial for interpreting internal and external processes affecting galaxies and identifying conditions for intense star formation in their early stages.Star Formation Rate

@@

where $\zeta$ accounts for mass loss, approximately $\zeta \approx 1.3$ (@kroupaConstraintsStarFormation2020).

@@

- **H\boldmath${\alpha}$ Emission**: Traces the number of ionizing photons, assuming full usage and re-emission. These photons mainly originate from hot, massive stars (O-type stars, \~10 Myr). Dust can significantly affect observations.

- **Infrared (IR) Flux**: Assumes a fixed fraction of stellar radiation is absorbed by dust, re-emitted at infrared wavelengths, and enhances total IR radiation.

@@

- **Far-Ultraviolet (FUV) Flux**: Mainly emitted by young, hot stars (e.g., B-type stars, \~100 Myr). Dust presence can also significantly affect observations.

@@

$$
\text{SFR} = \mathcal{K}_i\times L_i
$$ {#eq-sfr-spectra}

where $\mathcal{K}_i$ is a constant specific to each $L_i$ ($i =$ H$α$, IR, radio, FUV, X). In our analysis, we lack radio and X-ray data.

For FUV and H$α$ observations, which are affected by dust, corrections or models accounting for dust absorption are applied. Generally:

Following @madauCosmicStarFormation2014, this method may underestimate true SFRs, as different galaxy populations exhibit varying absorption mechanisms.

A more robust estimate of total SFR is:

where $L_{\text{FUV}}$ is uncorrected for absorption.

@@

1. **Exponential Decline (τ Model)**:

   where $\tau$ is the timescale and $t_{\text{sf}} = t - T_0$ is the star formation time.

2. **Delayed Exponential (τ Model)**:

   accounting for an initial rise in SFR before declining. Here, $\tau$ is the time to reach maximum SFR.

3. **Log-Normal Model**:

   with $\tau$ and $T_0$ as free parameters.

4. **Double Power Law**:

   where $\tau$ is the timescale, and $\alpha$, $\beta$ govern the SFR rise and fall.

Non-parametric models are also available, providing flexibility for galaxies with complex SFH patterns.

@@

# Hubble Law

## Redshift 
## Lookback time

@@

## Newton-Raphson
## Markov Chain Monte Carlo

@@

## Catalog Completeness
## Comparing the Catalogs
## The "special" case of the Star Formation Rates

@@

## According to pre-existing bibliography
### Problems with the method

>>>>

@@

### Hubble-de Vaucouleurs Classification

@@

## Star-Forming Regions

@@

Understanding SFH models is crucial for interpreting internal and external processes affecting galaxies and identifying conditions for intense star formation in their early stages.

@@

$$
\int_0^{t_{sf}} \text{SFR}(t)dt = \zeta M_*(t_{sf}),\ t_{sf}=\text{Time of Star Formation,} 
$$ {#eq-sfr_int}

where $\zeta$ accounts for mass loss during the Star Formation and is approximately $\zeta \approx 1.3$ (@kroupaConstraintsStarFormation2020).

@@

[^2]

[^2]:
    -   [@calzettiStarFormationRate2012], [@ASTR620GalaxiesFall]

-   **Hα** \boldmath$\alpha$ **Emission**: Young, hot, massive stars (O-type stars, \~10 Myr, \~20 $M_\odot$) produce a number of ionizing photons, which they ionize the surrounding hydrogen rich gas. The hydorgen undergoes recombination cascades which produce Balmer emission lines of $H\alpha$ (0.6563 $\mu m$) and $H\beta$ (0.4861 $\mu m$). Dust can significantly affect observations.

-   **Far-Ultraviolet (FUV) Flux**: Mainly emitted by young, hot stars ( B-type stars, \~100 Myr). Dust presence can also significantly affect observations.

-   **Infrared (IR) Flux**: The stars in a galaxy can heat up the dust in different ways, which then emits radiation in different parts of the IR spectrum. For example, young and massive, short-lived stars, emit UV radiation which then the heated dust emits in a wavelength of $\approx 60 \mu m$, whereas dust heated by UV-faint old or low-mass stars will emit at $\approx 100-150\mu m$. As a result, the total IR emission is age-agnostic and provides a more accurate approximation of the SFR because it accounts for contributions from both young and old stellar populations.

@@

@@

$$
\text{SFR}_i = \mathcal{K}_i\times L_i
$$ {#eq-sfr-spectra}

where $\mathcal{K}_i$ is a constant specific to each $L_i$ ($i =$ H$\alpha$, IR, radio, FUV, X). In our analysis, we lack radio and X-ray data.

Since the luminosities $L_{\text{FUV}}$ and $L_{\text{H}\alpha}$ originate from young stars and are highly sensitive to dust, we either directly observe stars unaffected by dust or use correction models to account for dust absorption. It is crucial to ensure that these models neither underestimate nor overestimate the luminosities by overlooking or double-counting the same sources.

Additionally, because these luminosities are emitted by similar stellar populations, we can reasonably expect the $SFR_{\text{FUV}}$ and $SFR_{\text{H}\alpha}$ to be approximately equal. As shown in the data from [@karachentsevSTARFORMATIONPROPERTIES2013] and supported by [@kroupaConstraintsStarFormation2020], a suitable approach for estimating the total SFR from FUV and H$\alpha$ observations is to calculate their average:

$$
SFR_{\text{FUV},H\alpha} = \text{mean}(SFR_{\text{FUV}}, SFR_{\text{H}\alpha}) 
$$ {#eq-SFR_comb}

where $L_{\text{FUV}}$ and $L_{\text{H}\alpha}$ are corrected for dust attenuation.

![Plot showing the linear relation $\log_{10}\text{SFR}_{FUV} = \log_{10}\text{SFR}_{H\alpha}$, as well as their distrubutions](figure/log_SFR_FUV_Ha.png){fig-align="center" width="500"}

According to [@madau2014], this method often underestimates the SFR, since different galaxy populations may systematically follow distinct absorption mechanisms depending on their characteristics.

Since $SFR_{\text{FUV}}$, based on the uncorrected $L_{FUV}$, represents the emission from unobstructed stellar populations, and $SFR_{TIR}$ isaccounts for dust-reprocessed light, a more accurate way to calculate the total SFR of a galaxy is:

$$
SFR_{\text{total}} = \mathcal{K}_{\text{FUV}} \cdot L_{\text{FUV}} + \mathcal{K}_{\text{IR}} \cdot L_{\text{IR}}
$$ {#eq-SFR_tot_madau}

Following the same reasoning as the previous formula, the total SFR can be expressed as:

$$
SFR_{\text{total}} = \text{mean}\left(\mathcal{K}_{\text{FUV}} \cdot L_{\text{FUV}}, \mathcal{K}_{\text{H}\alpha} \cdot L_{\text{H}\alpha}\right) + \mathcal{K}_{\text{IR}} \cdot L_{\text{IR}} \\
= SFR_{\text{FUV},H\alpha}+ SFR_{\text{IR}}
$$

where $L_{\text{FUV}}$ and $L_{\text{H}\alpha}$ are not corrected for dust absorption. However, since we do not have enough galaxies with both traces, we will use a different method of calculating the total SFR, which we will discuss later.

@@

-   **Exponential Decline (Tau Model):** The star formation rate (SFR) decreases exponentially over time, following the equation:

$$
  \text{SFR}(t) \propto e^{-t_{\text{sf}}/\tau}
$$

where $\tau$ is the timescale, $t_{\text{sf}} = t - T_0$ is the star formation time, $t$ is the age of the Universe, and $T_0$ is the time when star formation began.

-   **Delayed Exponential (Delayed Tau Model):** This model provides a more complex representation where the SFR initially increases, reaches a peak, and then declines exponentially over time. The equation for this model is:

$$
  \text{SFR}(t) \propto t_{\text{sf}} e^{-t_{\text{sf}}/\tau}
$$

This accounts for an initial growth phase followed by a decline. In this case, $\tau$ represents the time it takes for the galaxy to reach $\text{SFR}_{\text{max}}$.

-   **Log-Normal Distribution Model:** The SFR follows a normalized log-normal distribution, which can accurately model the star formation rate density ($\text{SFRD} = \text{SFR}/M_*$) in individual galaxies. The general form of the equation is:

$$
  \text{SFR}(t) \propto \frac{1}{\tau} \exp\left(-\frac{(\ln(t) - T_0)^2}{2\tau^2}\right)
$$

where $\tau$ and $T_0$ are free parameters of the distribution that lack physical significance, as the SFR does not necessarily peak at $t = e^{T_0}$.

-   **Double Power Law:** This model describes a scenario where the SFR rises and then falls sharply, useful for modeling galaxies experiencing rapid changes in star formation. The equation is:

$$
  \text{SFR}(t) \propto \left[\left(\frac{t}{\tau}\right)^\alpha + \left(\frac{t}{\tau}\right)^\beta\right]^{-1}
$$

where $\tau$ is the timescale and $\alpha$, $\beta$ are exponents that govern the rise and fall of the SFR.

```{bash}
gnuplot -e "
set terminal svg;
set output 'figure/combined_models.svg';
set xlabel 'Time (t)';
set ylabel 'SFR(t)';
set title 'Star Formation Rate Models';
set grid;
plot exp(-x/8) title 'Exponential Decline (tau=2)' with lines lw 2, \
     x*exp(-x/2) title 'Delayed Exponential (tau=2)' with lines lw 2, \
     (1/1)*exp(-((log(x)-2)**2)/(2*1**2)) title 'Log-Normal (tau=1, T0=2)' with lines lw 2, \
     1 / ((x/2)**2 + (x/2)**3) title 'Double Power Law (alpha=2, beta=3)' with lines lw 2;
"
```

```{r}
#| fig.cap="Star Formation Rate over the time of star formation, for different parametric models"
library(ggplot2)
library(dplyr)

# Define time values (positive times)
t <- seq(0.1, 30, length.out = 300)

# Define SFR models
exponential_decline <- function(t, tau) exp(-t / tau)
delayed_exponential <- function(t, tau) t * exp(-t / tau)
log_normal <- function(t, tau, T0) (1 / tau) * exp(-((log(t) - log(T0))^2) / (2 * tau^2))
double_power_law <- function(t, tau, alpha, beta) 1 / ((t / tau)^alpha + (t / tau)^(-beta))

# Generate data for each model
data <- data.frame(
  t = rep(t, 4),
  SFR = c(
    exponential_decline(t, tau = 2),
    delayed_exponential(t, tau = 2),
    log_normal(t, tau = 2, T0 = 12),
    double_power_law(t, tau = 2, alpha = 2, beta = 8)
  ),
  Model = rep(c("Exponential Decline", "Delayed Exponential", "Log-Normal", "Double Power Law"), each = length(t))
)

# Plot all models
ggplot(data, aes(x = t, y = SFR, color = Model)) +
  geom_line(size = 1) +
  labs(
    title = "Star Formation Rate Models",
    x = expression("Star Formation time (" * t[sf] * ") [Gyr]"),
    y = expression("SFR(" * t[sf] * ") ["* M['☉']/yr *"]" ),
    color = "Model"
  ) +
  scale_x_continuous(limits = c(0, 30)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

Additionally, there are non-parametric models, which do not follow a specific functional form to describe the star formation of a galaxy. These models are more flexible in adapting to galaxies with more complex star formation patterns.

@@

The Lilly-Madau plot is one of the most important plots in the field of galaxy evolution. It describes how the SFRD of the universe evolved over time, with observational data. to figure out

## Redshift and lookback time

According to Hubble–Lemaître law, all the galaxies are moving away from each other, at a speed proportional to their distance, due to the expansion of the universe.

$V = H_0\times d$, where $H_0 \approx 69.8 \text{ km/s/Mpc}$ is the Hubble constant and $d$ is the distance between the two galaxies.[^3]

[^3]: We also have non-Hubble motions $V = H_0\times d + V_0$, where $V_0$ is the peculiar velocity and it could, for example, be due to galaxy cluster dynamics. For the current explanation we are going to ignore it. This way the radial velocity $v$ is equal to $V$

Since we have galaxies with relative motions emitting light waves, we can observe the Doppler effect. Specifically, since the galaxies are moving away from each other, and thus from us also, we observe radiation with longer wavelenghts.

*Redshift* ($z$) is the doppler shift resulting from radial motion:

$$
z = \frac{\lambda_{observed}}{\lambda_{emitted}}-1
$$

In special relativity, $z$ is related to radial velocity $v$ by [@hoggDistanceMeasuresCosmology2000]

$$
1+z = \sqrt{\frac{1+v/c}{1-v/c}}
$$ {#eq-z-rel} For small $v/c$ we can rewrite @eq-z-rel, as:

$$
z \approx \frac{V}{c} = \frac{H_0\times d}{c}
$$

But, because light takes time to cover the distance $d$ between two galaxies, when the light finally reaches as, we will see the observed galaxy, as it was when the light was emitted, and not how it is at this moment. If we substitute time that it took the light to reach us over the distance, then we arrive at the relation [@longairGalaxyFormation1998]:

$$
t_\text{emitted} \propto z^{-3/2}
$$

*The lookback time* is the difference between the current age of the Universe and the age of the Universe when the light was emitted

$$
t_L = T_0-t_\text{emitted}
$$

@@

## Newton-Raphson

## Markov Chain Monte Carlo

@@

## Catalog Completeness

## Comparing the Catalogs

## The "special" case of the Star Formation Rates

@@

## According to pre-existing bibliography

### Problems with the method
